Portable Multi- and Many-Core Performance for Finite Difference Codes; Application to the Free-Surface Component of NEMO.

### 2017-06-20 docs/papers/nemolite2d_performance/nemolite2d_perf.tex
nemolite2d-paper: add methodology text and update GPU perf results

When executing an OpenMP-parallel program on a Non-Uniform Memory
Access (NUMA) compute node it becomes important to ensure that the
memory locations accessed by each thread are local to the hardware
core upon which it is executing. One way of doing this is to implement
a so-called 'first-touch policy' whereby memory addresses that will
generally be accessed by a given thread during program execution are
first initialised by that thread. This is simply achieved by using an
OpenMP-parallel loop to initialise newly-allocated arrays to some
value, \textit{e.g.} zero.

@@

Moving the kernels to execute on the GPU was achieved by using the
OpenACC \texttt{kernels} directive.

He also experimented with using CUDA directly within the
original form of NEMOLite2D in order to quantify any performance
penalty incurred by taking the {PS}y{KA}l/OpenACC approach.

>>>>

When executing an OpenMP-parallel program on a Non-Uniform Memory
Access (NUMA) compute node it becomes important to ensure that the
memory locations accessed by each thread are local to the hardware
core upon which it is executing. One way of doing this is to implement
a so-called `first-touch policy' whereby memory addresses that will
generally be accessed by a given thread during program execution are
first initialised by that thread. This is simply achieved by using an
OpenMP-parallel loop to initialise newly-allocated arrays to some
value, \textit{e.g.} zero.

@@

Moving the kernels to execute on the GPU was achieved by using the
OpenACC \texttt{kernels} directive in each Fortran routine containing
loops over grid points. (In the {PS}y{KA}l version this is just the
single {PS}y layer subroutine). This directive instructs the OpenACC
compiler to automatically create a GPU kernel for each loop nest it
encounters.

\subsubsection{Force parallelisation of Flather kernels}

The PGI compiler was unable to determine whether the loops applying
the Flather boundary condition in the $x$ and $y$ directions were safe
to parallelise. This information therefore had to be supplied by
inserting a \texttt{loop independent} OpenACC directive before each of
the two loops.

\subsubsection{Loop collapse}

When using the OpenACC \texttt{kernels} directive the compiler creates
GPU kernels by parallelising only the outer loop of each loop nest.
For the majority of loop nests in NEMOLite2D all of the iterations are
independent. It is therefore beneficial to expose the maximum amount
of parallelism to the GPU and we do this by instructing the compiler
to parallelise both levels (i.e. \texttt{ji} and \texttt{jj}) of a
loop nest. This is achieved using the OpenACC \texttt{loop collapse(2)}
directive.

\subsubsection{Shortloop}

As well as accounting for the bulk of the execution time, the Momentum
kernels ($u$ and $v$) are also relatively large.

\subsubsection{Kernel fusion}

\subsubsection{CUDA}

We also experimented with using CUDA directly within the original form
of NEMOLite2D in order to quantify any performance penalty incurred by
taking the {PS}y{KA}l/OpenACC approach.
